An efficient composition of bidirectional programs by tupling and lazy updates

### 2020-02-19 paper/4.xpg.tex
finish renaming + revise dummy

The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$. The second one tuples two results of $put$ and $get$ in the body of the if-expression. This is a trick since in some cases, the result of $pg$ may be undefined although the result is not undefined when evaluating $\getbxinline{Skip \ h}$. The last equality is relatively obvious.

    \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
%-{
    \textcolor{red}{
    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
        \tab \qtab \overset{2}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
            \qtab \tabs{2} \ \, (s_2, v_2) \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
            \tab \qtabs{2} ((s_1,s_2), (v_1,v_2))$\\
    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
        \tab \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{f_1 \ s}{v};\\
            \tab \qtabs{2} (f_2 \ s, v)$\\
    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
        \tab \qtab \overset{2}{=} (s, v) \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \tab \qtabs{2} (s, g_2 \ v)$
    }\\
%-}
    %+{
    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
        \tab \qtab \overset{2}{=} (s_1', v_1') \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
            \qtab \tabs{2} \ \, (s_2', v_2') \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
            \tab \qtabs{2} ((s_1',s_2'), (v_1',v_2'))$\\
    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{f_1 \ s}{v};\\
            \tab \qtabs{2} (f_2 \ s', v')$\\
    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \tab \qtabs{2} (s', g_2 \ v')$
    %+}

@@

    \noindent $\tabs{2} \pg{bx_1 \circ bx_2}{s}{v} \\
    \tabs{3} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tabs{3} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
            \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{construct\_dummy \ v}; \\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \tabs{2} \qtab (s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tabs{3} (s_3, v_2)$

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid.
%We can construct a $dummy$ from the source by \textsc{GetPut} law in some cases or require programmers to give a valid view.
Since there is no real view, this $dummy$ is necessary to pair with original source $s$ to form the input of $put \, [\![bx_1]\!]$. The last equality changes two. The first is a change from $dummy$ to an application $construct\_dummy \ v$. This means we need a way to construct $dummy$ from the original view $v$. Programmers can be required to give this $construct\_dummy$ function, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is the easiest way to meet our expectation. We set up that choice in our experiments but still use the $construct\_dummy$ function in the later definitions to provide a more general meanning.
%
The second is the change from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively. Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$. The substitution of $v_3$ by $v_3'$ is simply replacing the variable name since $v_3$ and $v_3'$ hold different results of $\getbx{bx_1}{s}$ and $\getbx{bx_1}{s_1}$ respectively. Because both variables are no longer used later, this substitution does not affect the outcome of the function.

@@

%-{
    \noindent \textcolor{red}{
    $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
        \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
        \qtab (f_2 \circ ks, kv, s, v)$
    }

    \noindent \textcolor{red}{
    $\cpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{s}{v} =\\
        \tab (ks, kv, s, v) \Leftarrow \cpg{bx}{ks}{g_1 \circ kv}{s}{g_1 \ v};\\
        \qtab (ks, g_2 \circ kv, ks, g_2 \ v)$
    }

    \noindent \textcolor{red}{
    $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
        \tab \textnormal{then} \ (ks, kv, s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
        \tab \textnormal{else} \ (ks, kv, s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
        \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks, kv, s', v')$
    }
%-}

    %+{
    \noindent $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
        \tab (ks', kv', s', v') \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
        \qtab (f_2 \circ ks', kv', s', v')$

@@

    \noindent $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
        \tab \textnormal{then} \ (ks', kv', s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
        \tab \textnormal{else} \ (ks', kv', s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
        \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks', kv', s', v')$
    %+}

    \noindent $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{construct\_dummy \ v};\\
        \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
            \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
\end{definition}

In the places where third and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
%Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$. The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, construct\_dummy \ v)$ where $s$ and $ks$ are corresponding to the source and the update over source. Since there is no real view here, we need to construct a dummy from the source same as $pg$. Then the continuation updating on this dummy should be initiated as an identity function. The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$. In the next assignment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input. It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$. After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.

@@

@@

\smallvspace
    $\tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{\lambda \_.s_0}{id}{id}{id}{s_0}{v_0};\\
        \tab \qtab (lks \ s; lkv \ v)$
\smallvspace

The beginning of accumulative functions $lks$ and $lkv$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$.

@@

%-{
    \noindent 
    \textcolor{red}{
        $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
        \tab \textnormal{if} \ h \ s = v \ \textnormal{then} \ (ks, kv, lks, lkv, s, v) \ \textnormal{else } \text{undefined}$
    }

    \noindent 
    \textcolor{red}{
        $\kpg{Replace}{ks}{kv}{lks}{lkv}{s}{v} = (kv, ks, lkv, lks, v, s)$
    }

    \noindent 
    \textcolor{red}{
        $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
        \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ lks}{fst \circ lkv}{s}{v};\\
        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ lks}{snd \circ lkv}{s}{v};\\
        \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
        \qtab \tab con \ (lks_1 \circ fst) \ (lks_2 \circ snd), \ con \ (lkv_1 \circ fst) \ (lkv_2 \circ snd), \\
        \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$
    }

    \noindent 
    \textcolor{red}{
        $\kpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{f_1 \circ ks}{kv}{f_1 \circ lks}{lkv}{s}{v};\\
        \qtab (f_2 \circ ks, kv, f_2 \circ lks, lkv, s, v)$
    }

    \noindent 
    \textcolor{red}{
        $\kpg{RearrV \ g_1 \ g_2 \ bx}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{ks}{g_1 \circ kv}{lks}{g_1 \circ lkv}{s}{v};\\
        \qtab (ks, g_2 \circ kv, lks, g_2 \circ lkv, s, v)$
    }

    \noindent 
    \textcolor{red}{
        $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab s \Leftarrow lks \ s; \quad v \Leftarrow lkv \ v; \quad lks \Leftarrow id; \quad lkv \Leftarrow id;\\
        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
        \tab \textnormal{then} \ (ks, kv, lks, lkv, s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{lks}{lkv}{s}{v}\\
        \tab \textnormal{else} \ (ks, kv, lks, lkv, s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{lks}{lkv}{s}{v}\\
        \tab \underline{\textnormal{fi}} \ cond_s \ (lks \ s') \ \&\& \ cond_{sv} \ s \ (lkv \ v'); \ \textnormal{return} \ (ks, kv, lks, lkv, s', v')$
    }

    \noindent 
    \textcolor{red}{
        $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{construct\_dummy \ v};\\
        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ lks_2, lkv_2, s_2, v_2)$
    }
%-}

    %+{
    \noindent $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
        \tab \textnormal{if} \ h \ es = ev \ \textnormal{then} \ (ks, kv, elks, elkv, es, ev) \ \textnormal{else } \text{undefined}$

@@

    \noindent $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
        \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst \circ elks}{fst \circ elkv}{es}{ev};\\
        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd \circ elks}{snd \circ elkv}{es}{ev};\\
        \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
        \qtab \tab con \ (lks_1 \circ fst) \ (lks_2 \circ snd), \ con \ (lkv_1 \circ fst) \ (lkv_2 \circ snd), \\
        \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$

@@

    \noindent $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v; \quad elks \Leftarrow id; \quad elkv \Leftarrow id;\\
        \tab \textnormal{if} \ cond_{sv} \ es \ ev \ \&\& \ cond_{s} \ es\\
        \tab \textnormal{then} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{elks}{elkv}{es}{ev}\\
        \tab \textnormal{else} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{elks}{elkv}{es}{ev}\\
        \tab \underline{\textnormal{fi}} \ cond_s \ (lks' \ s') \ \&\& \ cond_{sv} \ \textbf{es} \ (lkv' \ v'); \ \textnormal{return} \ (ks', kv', lks', lkv', s', v')$

    \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{construct\_dummy \ v};\\
        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ lks_2, lkv_2, s_2, v_2)$
    %+}
\end{definition}

@@

    \noindent $\xpg{bx}{s}{v} = \match{bx}\\
    \tab \vert \ bx_1 \circ bx_2 \to \\
        \tabs{2} (ks_1,kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{construct\_dummy \ v};\\
        \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{lkv_1 \ v_1}{v};\\
        \tabs{3} (ks_1 \ s_2, v_2)\\
    \tab \vert \ \_ \to \text{similar to } pg$
\end{definition}

>>>>

The first equality is simply based on the definitions of $pg$, $\putbxinline{Skip \ h}$ and $\getbxinline{Skip \ h}$. The second one tuples two results of $put$ and $get$ in the body of the if-expression. This is a trick since in some cases, the result of $pg$ may be undefined although the result is not undefined when evaluating $\getbxinline{Skip \ h}$. The last equality is a function application.

    \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
        \tab \qtab \overset{2}{=} (s_1', v_1') \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
            \qtab \tabs{2} \ \, (s_2', v_2') \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
            \tab \qtabs{2} ((s_1',s_2'), (v_1',v_2'))$\\
    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{f_1 \ s}{v};\\
            \tab \qtabs{2} (f_2 \ s', v')$\\
    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \tab \qtabs{2} (s', g_2 \ v')$

@@

    \noindent $\tabs{2} \pg{bx_1 \circ bx_2}{s}{v} \\
    \tabs{3} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tabs{3} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
            \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \tabs{2} \qtab (s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tabs{3} (s_3, v_2)$

The construction of $\pginline{bx_1 \circ bx_2}$ can be considered as the soul of the $pg$ function. The first two equalities comes from mentioned definitions and some basic transformations. The third one rewrites $v_1 \Leftarrow \getbx{bx_1}{s}$ into $(s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}$. This is possible when we consider $\getbx{bx_1}{s}$ as the second element of $\pg{bx_1}{s}{dummy}$ where $dummy$ is a special value that makes the $put \, [\![bx_1]\!]$ valid. Since there is no real view, this $dummy$ is necessary to pair with original source $s$ to form the input of $put \, [\![bx_1]\!]$. In general, $dummy$ depends on the source $s$, the view $v$ and/or the program $bx_1$. Programmers can be required to give a way to construct $dummy$, but it may be inessential for ill-typed systems where choosing $dummy$ as $v$ is one of the easiest ways to meet our expectation. That setting is used in our experiments. The last equality changes from $(s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}$ to $(s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}$, where $s$ and $v_3$ are replaced with $s_1$ and $v_3'$ respectively. Because $s_1$ is a source update of $\putbx{bx_1}{s}{dummy}$, so under the \textsc{PutPut} law, it is possible to substitute $s$ by $s_1$. The substitution of $v_3$ by $v_3'$ is simply replacing the variable name since $v_3$ and $v_3'$ hold different results of $\getbx{bx_1}{s}$ and $\getbx{bx_1}{s_1}$ respectively. Because both variables are no longer used later, this substitution does not affect the outcome of the function.

@@

    \noindent $\cpg{RearrS \ f_1 \ f_2 \ bx}{ks}{kv}{s}{v} =\\
        \tab (ks', kv', s', v') \Leftarrow \cpg{bx}{f_1 \circ ks}{kv}{f_1 \ s}{v};\\
        \qtab (f_2 \circ ks', kv', s', v')$

@@

    \noindent $\cpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{s}{v} =\\
        \tab \textnormal{if} \ cond_{sv} \ s \ v \ \&\& \ cond_{s} \ s\\
        \tab \textnormal{then} \ (ks', kv', s', v') \Leftarrow \cpg{bx_1}{ks}{kv}{s}{v}\\
        \tab \textnormal{else} \ (ks', kv', s', v') \Leftarrow \cpg{bx_2}{ks}{kv}{s}{v}\\
        \tab \textnormal{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \ \textnormal{return} \ (ks', kv', s', v')$

    \noindent $\cpg{bx_1 \circ bx_2}{ks}{kv}{s}{v} =\\
        \tab (ks_1, kv_1, \underline{s_1}, v_1) \Leftarrow \cpg{bx_1}{ks}{id}{s}{dummy};\\
        \tab (ks_2, kv_2, s_2, v_2) \Leftarrow \cpg{bx_2}{kv_1}{kv}{v_1}{v};\\
            \qtab (ks_1 \circ ks_2, kv_2,  ks_1 \ s_2, v_2)$
\end{definition}

In the places where third and/or fourth argument ($s$ and $v$) are updated by applications, the computations are also accumulated in $ks$ and/or $kv$.
%Apart from the last construction, the others are quite similar to the corresponding ones of $pg$, but have some updates over the source and the view on accumulative functions $ks$ and $kv$ respectively.
Thanks to these accumulations, there are only two $cpg$ calls in $\cpginline{bx_1 \circ bx_2}$. The first call $\cpginline{bx_1}$ requires parameter $(ks, id, s, dummy)$ where $s$ and $ks$ are corresponding to the source and the update over source. Since there is no real view here, we need a dummy same as $pg$. Then the continuation updating on this dummy should be initiated as an identity function. The first $cpg$ call is assigned to a 4-tuple $(ks_1, kv_1, s_1, v_1)$. In the next assignment, a 4-tuple $(ks_2, kv_2, s_2, v_2)$ is assigned by the second $cpg$ call which uses the input as $(kv_1, kv, v_1, v)$ where $kv_1$ and $v_1$ are obtained from the result of the first assignment, and $kv$ and $v$ come from the input. It is relatively similar to the second $pg$ call assignment in $\pginline{bx_1 \circ bx_2}$. After two $cpg$ calls, a function application, $ks_1 \ s_2$, is used to produce the updated source instead of calling recursively one more time like in $\pginline{bx_1 \circ bx_2}$.

@@

The result pair $(s, v)$ obtained from $cpg$ as above should be same with the result of $\pg{bx}{s_0}{v_0}$. More generally, we have the following relationship:

$\tab \cpg{bx}{ks}{kv}{s}{v} = \pg{bx}{ks \ s}{kv \ v}$

@@

\smallvspace
    $\tab (ks, kv, lks, lkv, s, v) \Leftarrow \kpg{bx}{\lambda \_.s_0}{id}{id}{id}{s_0}{v_0};\\
        \tab \qtab (lks \ s, lkv \ v)$
\smallvspace

The beginning of accumulative functions $lks$ and $lkv$ are set as identity function, while $ks$ and $kv$ are initiated as the same with the corresponding ones in $cpg$. The relationship among $kpg$, $cpg$ and $pg$ can be shown as follows:

$\tab \kpg{bx}{ks}{kv}{lks}{lkv}{s}{v} = \cpg{bx}{ks \circ lks}{kv \circ lkv}{s}{v}\\
    \tabs{5} = \pg{bx}{ks \ (lks \ s)}{kv \ (lkv \ v)}$

@@

    %+{
    \noindent $\kpg{Skip \ h}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v;\\
        \tab \textnormal{if} \ h \ es = ev \ \textnormal{then} \ (ks, kv, id, id, es, ev) \ \textnormal{else } \text{undefined}$

@@

    \noindent $\kpg{\product{bx_1}{bx_2}}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v;\\
        \tab (ks_1, kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \\
            \tabs{5} \kpg{bx_1}{fst \circ ks}{fst \circ kv}{fst}{fst}{es}{ev};\\
        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \\ 
            \tabs{5} \kpg{bx_2}{snd \circ ks}{snd \circ kv}{snd}{snd}{es}{ev};\\
        \qtab ( \ con \ ks_1 \ ks_2, \ con \ kv_1 \ kv_2, \\
        \qtab \tab con \ (lks_1 \circ fst) \ (lks_2 \circ snd), \ con \ (lkv_1 \circ fst) \ (lkv_2 \circ snd), \\
        \qtab \tab (s_1, s_2), \ (v_1,v_2) \ )$

@@

    \noindent $\kpg{Case \ cond_{sv} \ cond_{s} \ bx_1 \ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab es \Leftarrow lks \ s; \quad ev \Leftarrow lkv \ v;\\
        \tab \textnormal{if} \ cond_{sv} \ es \ ev \ \&\& \ cond_{s} \ es\\
        \tab \textnormal{then} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_1}{ks}{kv}{id}{id}{es}{ev}\\
        \tab \textnormal{else} \ (ks', kv', lks', lkv', s', v') \Leftarrow \kpg{bx_2}{ks}{kv}{id}{id}{es}{ev}\\
        \tab \underline{\textnormal{fi}} \ cond_s \ (lks' \ s') \ \&\& \ cond_{sv} \ es \ (lkv' \ v'); \ \textnormal{return} \ (ks', kv', lks', lkv', s', v')$

    \noindent $\kpg{bx_1 \circ bx_2}{ks}{kv}{lks}{lkv}{s}{v} =\\
        \tab (ks_1, kv_1, \underline{lks_1}, lkv_1, \underline{s_1}, v_1) \Leftarrow \kpg{bx_1}{ks}{id}{lks}{id}{s}{dummy};\\
        \tab (ks_2, kv_2, lks_2, lkv_2, s_2, v_2) \Leftarrow \kpg{bx_2}{kv_1}{kv}{lkv_1}{lkv}{v_1}{v};\\
        \qtab (ks_1 \circ ks_2, kv_2, ks_1 \circ lks_2, lkv_2, s_2, v_2)$
    %+}
\end{definition}

@@

    \noindent $\xpg{bx}{s}{v} = \match{bx}\\
    \tab \vert \ bx_1 \circ bx_2 \to \\
        \tabs{2} (ks_1,kv_1, lks_1, lkv_1, s_1, v_1) \Leftarrow \kpg{bx_1}{\lambda \_.s}{id}{id}{id}{s}{dummy};\\
        \tabs{2} (s_2, v_2) \Leftarrow \xpg{bx_2}{lkv_1 \ v_1}{v};\\
        \tabs{3} (ks_1 \ s_2, v_2)\\
    \tab \vert \ \_ \to \text{similar to } pg$
\end{definition}

### 2020-02-19 paper/5.experiments.tex
finish renaming + revise dummy

We have fully implemented and tested all methods% 
%
\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default. Our target language is untyped.}$^,$%HACK!
%
\footnote{Anonymous \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
%
described in the previous sections.
Instead of constructing some dummies by $construct\_dummy$, which are necessary for $pg$, $cpg$, $kpg$ and $xpg$, we replace them by the current updated views. This helps a program in the $put$ direction valid.

>>>>

We have fully implemented and tested all methods% 
%
\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.}$^,$%HACK!
%
\footnote{Anonymous \url{https://mega.nz/\#F!ec8y2ITa!tLOHKkHgs_7fjSI8m-O26Q}}
%
described in the previous sections. Our target language is untyped. Some dummies used for $pg$, $cpg$, $kpg$ and $xpg$ are replaced with the current updated views. This helps a program in the $put$ direction valid.
